---
title: "Name: Simon Chen"
author: 
- |
    | Student number: 17016335
date: "`r format(Sys.time(), '%X, %d %B, %Y')`"
output: html_document
---
# Originality declaration  

I, [Simon Tianzuo Chen], confirm that the work presented in this assessment is my own. Where information has been derived from other sources, I confirm that this has been indicated in the work.

date: `r format(Sys.time(), '%d %B, %Y')`

# Start your response here


## Initial project scope

The following data is collected from The City of New York's Open Data hub for the purpose of analysing and understanding the possible factors that impact the likelihood of eviction. 

To perform these analysis, appropriate data processing and analysis methods will be utilised, with an overview report being prepared that will showcase some of the patterns that arise from our analysis. 

To begin, we will install some packages that will be useful for our analysis
```{r}
# Installing Packages
library(sf)
library(tmap)
library(sp)
library(tmaptools)
library(tidyverse)
library(janitor)
library(here)
```

From there, we'll load in our datafiles, a shapefile of New York and eviction data
```{r}
# load in the data 
NYC <- read_sf(here::here("Community Districts", "geo_export_e06d52f8-300a-4a76-9d41-23cdb61499cb.shp"))
evictions <- read_csv(here::here("Evictions.csv"))
qtm(NYC)
```